---
title: "Visualizations"
output: html_notebook
---

## Introduction
This notebook produces maps from the data processed in prior notebooks.

## Dependencies
These are the packages neccessary for making these maps.
```{r dependencies}
library(sf)
library(dplyr)
library(ggplot2)
library(viridis)
library(gateway)
```

## Load Data
These are the data we will be mapping:
```{r load data}
m14 <- st_read("../data/Spatial/grid/full_14.shp") %>% mutate(year = 2014)
m16 <- st_read("../data/Spatial/grid/full_16.shp") %>% mutate(year = 2016)
m17 <- st_read("../data/Spatial/grid/full_17.shp") %>% mutate(year = 2017)

m_all <- rbind(m14, m16, m17)
# bg data
bg = st_union(gateway::gw_get_data("Grids", "sf"))
```

## Making a Choropleth
These maps dont have labels because they are used on the poster for this work. Therefore they are saved with intuitive names and in a vector graphics format.

```{r ggplot csb}
# csb map
ggplot() +
  geom_sf(data = bg) +
  geom_sf(data = m_all, mapping = aes(fill = csb_dns)) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_line(color = NA),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "transparent", color = NA),
        legend.background = element_rect(fill = "transparent", color = NA)
        ) +
  scale_fill_viridis(name = "CSB Calls\nper 1000\nResidents", guide = guide_colorbar(ticks = F), direction = -1) +
  facet_grid(. ~ year)

ggsave("../results/csb_year.pdf", dpi = 500, bg = "transparent")
```

```{r ggplot voter}
# Voter map
ggplot() +
  geom_sf(data = bg) +
  geom_sf(data = m_all, mapping = aes(fill = turnout)) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_line(color = NA),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "transparent", color = NA),
        legend.background = element_rect(fill = "transparent", color = NA)) +
  scale_fill_viridis(option = "plasma", name = "% Voter\nTurnout", guide = guide_colorbar(ticks = F), direction = -1) +
  facet_grid(. ~ year)


ggsave("../results/voter_year.pdf", dpi = 500, bg = "transparent")
```


